\TheTitle

### 2016-04-22 linvis.tex
Bubble Tree, and editing
Making a bunch of edits.

        With an average of more than 900 top-level merges into the Linux kernel per
        release, many containing hundreds of commits and some containing thousands,
        maintenance of older versions of the kernel becomes nearly impossible.

        Various commercial products, such as the Android platform, run older
        versions of the kernel. Due to security, performance, and changing hardware
        needs, maintainers must understand how updates made to the current version
        of the kernel fit into older versions in order to make the necessary
        patches.

@@

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects that
are the scale of the kernel. Tools like Gitk provide users with the full view of the
DAG, showing all of the merges and commits. This information is useful in smaller
projects, showing the relation between various branches. The DAG becomes a tangled
mess (\ref{fig:gitk}) of commits, merges, and the links between them with large
modular projects, like the Linux kernel, making it difficult to derive an
explanation of the changes made. In other cases, the kernel is simply too large for
the system to generate a visualization for. Github provides a DAG view for most
projects, but is unable to display the visualization for projects as large as the
Linux kernel.

@@

Git log, Gitk, and other similar tools only store the name of the author, authored
date of a commit. A commit may be merged into the project at any time, regardless of
when the commit was authored, making date-range queries meaningless. Performing a
date-range query using these tools over the span of a release will contain results
that are for previous versions of the kernel or may not contain all of the results
for that kernel if patches were merged after the release date.

German et al. recorded the metadata of every commit and merge in the Linux git
repository starting in 2011 \cite{German}, including chronological data that is
not stored by git. We are able to use this information to generate sets of
trees instead of a DAG. Each tree is rooted by the top-level merge, performed
by Linus, which merges the branches or a set of commits into the kernel.

In this paper, we present our design decisions behind a web-based
tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
model instead of a DAG. Our tool provides information about the location of a
given commit in the respective merge tree, the files edited, the modules
edited, and the commit message for a given commit or merge. The tool allows
users to apply various filters, including the release version, along with a
keyword or phrase from the log preview, the name of the author, or the commit
id. The user can request all the top-level merges containing a commit or merge
that matches the query, or all commits and merges that match the query.

Our core model provides advantages to the DAG model used by other tools.
Breaking the commits and merges by merge tree removes commits and merges of
other components of the kernel from the visualization, providing a clearer
picture of what is happening in a given merge and providing additional
support for edited file information. Furthermore, we are able to understand the
changes being made, in the order they are brought into the kernel.

All top-level merges to a given version of the kernel are merged between the
start and release date. The tree construction allows us to find all commits that are
related to a given release of the kernel regardless of the authorship date, allowing
our system to provide better mechanisms for performing date-range queries.

\section{Related Work}
To our knowledge, we know of no repository visualization systems that builds a tree
from the merge. CVS and svn construct trees; however, the root node for these trees
is the branch point. This may be due to the fact that more information is required
to generate the trees than what is stored by git. A server must sit and monitor the
repository as it changes over time. Many of the visualization tools are designed for
a general-purpose approach, where any repository is directly usable without
additional information. Polling for this additional information makes the
application specific to a single repository or project.

\subsection{Gitk}
Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
the central organization model. For smaller projects, this model is sufficient
for providing meaningful information, but in larger projects like the Linux
kernel, the branches become too tangled and too deep to derive any meaning from
the DAG. The DAG model allows users to go to the next commit in the chain, go
to the previous commit in the chain, or jump to the parent merge. This suggest
that Gitk and similar tools are designed for users working from a given commit or
leaf-node toward the root.  With this model, it is difficult for users to see what
other commits are involved with a given merge. Gitk only provides file information
on commits, as it is unable to quickly aggregate the files and commits that are part
of a merge. It is up to the user to remember what files were edited in each of the
commits involved in a given merge. In some of the larger merges, some containing
over 1500 edited files, this task become nearly impossible.  % cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}), are
unable to provide a visualization for projects that are as large as the kernel.

\begin{figure}[h!]
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project}
        \label{fig:gitfail}
\end{figure}

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap visualization
tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
pictured in figure \ref{fig:treemap}, for
displaying the file structure of a given version of the kernel. Treemaps are good
for visualizing the topology of shallow, wide trees, making it a good candidate for
showing file systems. Treemaps are limited in what information can be displayed. The
information within the treemap does not provide clear insight on what changes were
made to the files in the given release, only providing information on where the file
is located within the file system.  File systems and the merge tree of the kernel
share many similarities, in both cases, they represent trees that are wide and
shallow, suggesting that a treemap could potentially work to visualize the commit
and merge structure of the kernel.

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
        \caption{Treemap of Linux 2.5.33 file structure}
        \label{fig:treemap}
\end{figure}

\subsection{List viewer}
Moghaddam designed a web-based visualization
tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
displaying the chain of merges into the kernel. There is no information on what
commits are within a given merge, or what files were edited. The visualization
tool provides information about who authored the commit and when it was
authored. We can see the visualization tool in figure \ref{fig:listviewer}
displaying a part of the kernel commit information, from commit ``26b23ac''
according to the tool.

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitvis.png}
        \caption{List Viewer}
        \label{fig:listviewer}
\end{figure}

\section{Design}
The goal is to build a tool that makes navigation of the kernel commit
information simpler, while providing an explanation of the changes occurring at
each commit and merge. Part of making the information usable is to narrow the
results down into only the information that is relevant to what the user is
looking for. We designed the tool with two use-cases in mind, though a user may
switch roles as they work.

\textit{Use-case 1: top-to-bottom approach}\\
These are users that are maintaining a section of the kernel and would like to
pick a merge and merge it directly into their current repository. This is
useful for minimizing required implementation on the side of the maintainer,
allowing them to see what changes are made higher up in the kernel and bundle
multiple changes together. We must provide them with methods of navigating from
a more general set of merges toward a specific set of merges and commits.

\textit{Use-case 2: bottom-to-top approach}\\
These are users that know what merge or commit they are working with and would
like to see what else is happening around their commit. This is useful to see
how changes to a given file will break other changes made in other commits.
This is primarily for maintainers that have to perform some implementation work
in order to correctly merge the patches from the future. These users may also
know what module(s) they are dealing with, and are looking for modules that are
relevant to the area they are working on. We must provide them methods of
navigating from a specific commit to a more general merge, allowing them to see
a bigger picture.

\subsection{Core Model}
Each top-level merge into the kernel is the root-node to a tree. Each merge
may contain zero or more children made of other merges or commits. Commits have
no children, and therefore are always leaf nodes.  The commits contain the
information about the files that are edited along with how many lines are added
and removed from each file. Commits also store the module of the kernel that
they are working with.

\subsection{Command-line Shell}
Our initial intuition was to work with the commits and merges as trees. We use
the merges as directories, and commits as files, like in a file structure.
There are many similarities between a file system and the structure of the
commit information. The primary similarity is that they are both trees, and
both trees are generally fairly shallow and wide. File systems have been
around for quite some time, and there are many tools for visualizing and
working withe the file structure. The first tool was the command-line. We built
a small command-line shell (\ref{fig:shell}) that allows navigation through
the commit and merges instead of files and directory. The shell design
permitted a user to move through the merges like they would through
directories with the \textit{cd} and \textit{ls} commands. Users can use the
\textit{cat} command to see the preview of the log, or the \textit{more}
command to see the preview and what files were edited in the commit and how
many lines were added and removed from each file.

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/shell.png}
        \caption{Shell Design}
        \label{fig:shell}
\end{figure}

There are two issues with this design. One issue with is that a user will
always move from the top-down without any additional filtering. This only
serves one of our use-cases and can be very overwhelming. The second issue
being that it presents the user with the commit hashes, providing no
explanation of what the commit or merge may contain.

\subsection{Web-based Tool}
There are multiple reasons for building the system into a web-based tool.
Putting the tool on the web enables users to use the system without having to
install any software, making the system more accessible, more easily
maintained, and platform independent. A web-based tool can utilize more
interactive means of navigation, and provide a better explanation of what each
commit or merge may contain than the command-line shell.

Our implementation of the web-based tool provides many mechanisms to enable
users to better navigate and understand changes in the kernel. These mechanisms
include searching, the commit messages, the files edited, the modules edited,
and two types of tree viewers.

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View}
        \label{fig:search}
\end{figure}

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results.png}
        \caption{Search Results}
        \label{fig:results}
\end{figure}

In the search results, the user is presented with the one-line log message
preview, the author's name and email, the date the commit was authored, and the
date the commit was last committed.  The commit date and author date will
usually be the same, however, if the commit has been rebased, the commit date
will be different than the authored date, and will always be after the authored
date. The first and third entries in figure \ref{fig:results} have been rebased
in the past, though we are unable to highlight rebases at this time.

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/message_view.png}
        \caption{Commit Message}
        \label{fig:message}
\end{figure}

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/file_view.png}
        \caption{Commit Files}
        \label{fig:files}
\end{figure}

The second tab is the files tab. This tab provides information on what files
have been edited, how many lines were added, and how many lines were removed in
a given commit. This functionality is similar to the other tools available. Our
tree-based design model allows us to extend this functionality to merges, which
the other tools are unable to show. When a file is edited multiple times from
different commits within a merge, we simply add the added and removed lines to
the values already stored.\newpage

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules.png}
        \caption{Commit Modules}
        \label{fig:modules}
\end{figure}

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
        \caption{Commit Tree View}
        \label{fig:tree}
\end{figure}

The third tab shows the tree view (\ref{fig:tree}). The tree view is what makes
this tool unique to the other tools, providing a full view of the entire merge
tree. It originally centres at the current commit or merge, but allows a
user to inspect the other commits and merges surrounding it. A user may then
navigate to any of the commits by simply double-clicking on it.

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
        \caption{Tree zoomed out}
        \label{fig:zoomed_tree}
\end{figure}

Some of the merge trees are very large (\ref{fig:zoomed_tree}), containing
thousands of commits and merges. Clicking on subtrees will cause them to
minimize, only showing parts of the tree that are of interest to the user.
Furthermore, the user can zoom the tree by scrolling the mouse wheel to see
more of the topology of the tree, or to see the details of a given part of the
tree.

The last tab contains another type of tree view. This tree is in the form of
nested lists, and is designed to more closely model the tree hierarchy view of
file browsers. This tree only contains the commits and merges that are within
the subtrees of the current merge. A commit will never have any items in this
tab, as it is a leaf node, so it cannot contain subtrees. To accompany the
tree, we include breadcrumbs at the top. The last item in the breadcrumb list
is the current commit, the next item is the parent of the current commit, and
the first item is the top-level merge into the kernel.

\section{Implementation Details}
We built a web-based tool to provide information about the kernel.
We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
as the framework for generating the dynamic content.

@@

Our tool provides mechanisms and visualizations that other tools are unable to
produce. Continuous mining of the repository\cite{German} provides an
additional dimension to the dataset that would otherwise be lost, enabling us
to reliably build trees from the information we have.

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
        \caption{Merge Dag View}
        \label{fig:dag_view}
\end{figure}

\begin{figure}[h]
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.png}
        \caption{Merge Tree View}
        \label{fig:tree_view}
\end{figure}

@@

\subsection{Performance Evaluation}
Originally, we were performing all of our tree-navigation, file information
gathering, and module information gathering using recursive SQL queries.
Recursive queries have terrible performance, with the average response time
from the database being around 3 seconds for a single operation. We found that
it was far more response time-efficient to load an entire merge tree directly
into memory, then iterate over the objects returned to build the tree
structures, than it was to have the dbms perform the tree traversals, building
the specific trees for a commit. At this time, we have re-implemented the
breadcrumbs, the files, and the main tree views to enhance the response times.
We will discuss some We will discuss the changes made.

The breadcrumbs are generated on page-load, so they must be generated quickly.
We began with a single recursive SQL query, but was unable get the results in
the desired time, causing all results pages to render slowly. We replaced the
recursive query with a single query that would grab the entire merge tree,
rooted at the top-level merge.

The items returned from the database contained the commit id, the commit id of
the merge with the Linux kernel, and the next merge in the path to the
top-level merge.

We placed each item in the results into a dictionary using the commit id as the
key. Using the following method, we build the dictionary into a list of commit
IDs for the breadcrumbs.

\begin{verbatim}
current = cid
breadcrumbs = [cid]
while current is not top-level merge:
        breadcrumbs.append(current.next)
        current = current.next
return reverse(breacrumbs)
\end{verbatim}

This maintains the O(n) runtime for generating the breadcrumbs, while using an
iterative method for traversing the tree and dropping the response time from
around 3 seconds to at most 500ms.

We use similar methods to generate the file information and the tree. Both are
generated asynchronously, allowing us to use spinners to help improve the user
experience. We are using AJAX requests to generate the asynchronous requests,
are are having very good results.\newpage

\subsection{Extendability}
Due to how the asynchronous requests are performed, it is possible to query the
server for the tree and file information directly, allowing users to build
their own front-ends for working with our tree-based system.

The tree and file information is accessible through
http://li.turingmachine.org/data/tree/JSON/<commit id> and
http://li.turingmachine.org/data/files/JSON/<commit id>, respectively.
The tree is generated as many nested javascript objects. The tree response does
not contain any notion of the current commit and will only return the entire
merge tree.

The tree responses are generated as a single object. This object is the root
node of the tree, and may contain a list of children in the ``children'' field.
The full response format for the tree looks like:

\begin{itemize}
        \item ``cid'': string, commit id
        \item ``name'': string, One-line log preview
        \item ``author'': string, Author's name
        \item ``mlinus'': string, commit id of top-level merge
        \item ``mnext'': string, commit id of the next merge on path to
                top-level
        \item ``children'': list, list of items with the same fields. Null if
                no children
\end{itemize}

The file responses contain only the files that the selected merge or commit
works with. The response is a single object in the form of a tree. This object
is the root of the tree, and represents the current merge or commit. If the
current position in the tree is an inner node, the response will contain the
child nodes in the ``children'' field, otherwise ``children'' will be null. The
full response format for the file response looks like:

\begin{itemize}
        \item ``cid'': string, commit id
        \item ``files'': list, list of tulples in the format
                \begin{enumerate}
                        \item Filename: string
                        \item Lines added: unsigned integer
                        \item Lines removed: unsigned integer
                \end{enumerate}
        \item ``mnext'': string, the commit id of the next merge toward the
                top-level merge
        \item ``children'': list, List of all child objects with the same
                fields. Null if no children.
\end{itemize}

\section{Future Work}
This project was implemented over four months, interleaved with other course
projects and assignments. There are still many areas that could use
improvements before the full potential of our model can be realized.
Here, we outline various areas that still need more attention.

\subsection{Files}
There is currently no functionality surrounding searching by filename. Use-case
2 may know what file they are editing and try to determine how it may work with
the other commits and merges in the module. It is possible that a third
use-case may arise, where the user wants to determine all commits that effect a
given file. In both cases, a bottom-to-top approach is used. With this
functionality, implementing the top-to-bottom search feature becomes trivial.

@@

\subsection{Evaluation}
At this time, we have no evidence that our tool is able to improve the
work-flow of maintainers. We believe that the tool is able to improve the
work-flow and performance of maintainers because it provides cleaner mechanisms
of visualizing and presenting the commit information. It is able to provide
more relevant information, while removing information that is irrelevant to a
given module or set of merges.

>>>>

        With an average of more than 900 top-level merges into the Linux kernel per
        release, many containing hundreds of commits and some containing thousands,
        maintenance of older versions of the kernel becomes nearly impossible.
        Various commercial products, such as the Android platform, run older
        versions of the kernel. Due to security, performance, and changing hardware
        needs, maintainers must understand how updates made to the current version
        of the kernel fit into older versions in order to make the necessary
        patches.

@@

Maintainers of older versions of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects at
the scale of the kernel. Tools like Gitk use the DAG model, showing all commits and
merges in chronological order to when the commit was authored. This representation
works in smaller projects, enabling users to see when changes are made, when these
changes are merged, and how each branch is interacting. In large modular projects,
like the Linux kernel, the DAG becomes a mess of merges and commits (figure
\ref{fig:gitk}), losing its visual meaning. In some cases, the Linux kernel is
simply too large for the system to generate a visualization. Github provides a DAG
view for many projects, but is unable to display the visualization for projects as
large as the Linux kernel (figure \ref{fig:gitfail}).

@@

Git cannot store the date that a commit was merged into a branch due to the rebase
operation.  When a rebase occurs, several changes could be made to the commit. It
may be squashed with other commits, combining the effects of several commits into a
single commit. It may be dropped altogether, in which case the commit is no-longer
merged.  It may be that the entire branch has simply been re-rooted to another
position on the DAG. Due to this, tools are unable to determine when a commit is
merged. A user may decide that it is important to be able to see only the commits
contributing to a given release, that is commits that were merged with the
Linus-branch of the kernel between two given dates. This operation is not possible
without the collection of additional data that is not provided by git. Git only
enables users to see when commit was authored, not when it was merged into a branch.

German et al. recorded the metadata of every commit and merge in the Linux git
repository starting in 2011 \cite{German2015}, including chronological data that is
not stored by git. We are able to use this information to determine when a commit
has been merged into the kernel, regardless of rebases preceding and proceeding that
point in time. This information is then converted by a heuristic method, described
by German, to generate sets of trees, which we use as our core model instead of
using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
merges the branches or a set of commits into the kernel.

In this paper, we present our design decisions behind a web-based
tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based model
instead of a DAG. Our tool provides information about the location of a given commit
in the respective merge tree, the files edited, the modules edited, and the commit
message for a given commit or merge. The tool allows users to apply various filters,
including the release version, along with a keyword or phrase from the log preview,
the name of the author, or the commit id. The user can request all the root-level
merges containing a commit or merge that matches the query, or all commits and
merges that match the query.

Using the tree provides many advantages to the DAG model, but there are two
drawbacks. The first drawback is that a server must query the state of the master
repository since time information is not stored by git. The second drawback is that
the tree construction is a heuristic approach, and in some cases may break.

Our core model provides advantages to the DAG model used by other tools. By
separating and focusing on sub-trees from the whole repository, we are able to
filter out commits and merges that do not pertain to a component of the kernel. This
provides a clearer picture of what is happening in a given merge. The second benefit
to the tree model is that it enables us to automatically aggregate metadata.

In addition to this, the model provides us with a mechanism of performing date-range
queries with meaningful results. Using the tree built by a continuous mining system
enables us a mechanism to perform date-range queries with meaningful results.

\section{Related Work}

To our knowledge, we know of no git repository visualization systems that builds a
tree from the merge. This may be because more information is required to generate
the trees than what is stored by git. Git uses a DAG model, which enables
history-altering actions like rebases to occur, whereas older tools like CVS and svn
construct trees\cite{CVS2008} but do not work with history-altering actions.

A server must sit and monitor the git repository as it changes over time to handle
changes to history. Many of the visualization tools are designed for a
general-purpose approach, where any repository is directly usable without additional
information. Polling for this additional information makes the application specific
to a repository or project.

\subsection{Other Visualizers}

Gitk, Gitg, and other similar git repository visualizing tools use the DAG as the
central organization model. This model is sufficient for smaller projects, providing
meaningful information; however, in larger projects, like the Linux kernel, the
branches become too tangled and too deep to derive any meaning from the DAG. Users
can interact with the DAG by going to the previous commit in the chain or jumping to
the parent merge. This suggests that Gitk and similar tools are designed for users
working from a given commit or leaf-node toward the root. With this model, it is
difficult for users to see what other commits are involved with a given merge.

Furthermore, the DAG model is only able to provide file-level information on a
commit level. It is unable to quickly aggregate the file data over all commits that
are part of a merge. If a user would like to view what files were edited at a merge,
they must manually aggregate this information by hand. While some merges are
relatively small, containing less than 10 commits, other merges are large,
containing over 1500 edited files, as in the case of commit \textit{73287a4}. This
issue is important for authorship and licensing. A merge can only show one author,
the contributor making the merge, but does not show who wrote the commits contained
within that merge.

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project}
        \label{fig:gitfail}
\end{figure}

@@

\subsection{Treemap visualizer}

University of Maryland presented a treemap visualization
tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
pictured in figure \ref{fig:treemap}, for
displaying the file structure of a given version of the kernel. Treemaps are good
for visualizing the topology of shallow, wide trees, making it a good candidate for
showing file systems. Treemaps are limited in what information can be displayed. The
information within the treemap does not provide clear insight on what changes were
made to the files in the given release, only providing information on where the file
is located within the file system. File systems and the merge tree of the kernel
share many similarities, in both cases, they represent trees that are wide and
shallow, suggesting that a treemap could potentially work to visualize the commit
and merge structure of the kernel.

@@

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/kernel-files.png}
        \caption{Treemap of Linux 2.5.33 file structure}
        \label{fig:treemap}
\end{figure}

\subsection{List viewer}

Moghaddam designed a web-based visualization
tool\footnote{\url{http://web.uvic.ca/~arasbm/gitVisualizations/linuxRGraph.html}}
displaying the chain of merges into the kernel. There is no information on what
commits are within a given merge, or what files were edited. The visualization
tool provides information about who authored the commit and when it was
authored. We can see the visualization tool in figure \ref{fig:listviewer}
displaying a part of the kernel commit information, from commit ``26b23ac''
according to the tool.

@@

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitvis.png}
        \caption{List Viewer}
        \label{fig:listviewer}
\end{figure}

\section{Methodologies}
\evan{Correct section title? Other than just trying stuff, what were the
methodologies?}

The goal is to build a tool that makes navigation of the kernel commit
information simpler, while providing an explanation of the changes occurring at
each commit and merge. We designed the tool with two use-cases in mind, though a
sumer may switch between the cases as they work.

\textit{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
These are users that are maintaining a section of the kernel and would like to pick
a merge and merge it directly into their current repository, which is useful for
reducing the amount of re-implementation work. For them, it is
important to have the ability to aggregate files and modules that are being changed
somewhere in the merge. It is also important for them to be able to navigate from
the root toward the largest merge that contains only the commits that modify the
modules that they are working with.

\textit{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
These are users that start with a known merge or commit and would like to see
what other changes are being made in commits that are in the same merge. This is
useful to see how changes in a given file will break other changes made in the
surrounding commits in the same merge. This is primarily for maintainers that need
to perform some implementation work in order to correctly merge patches from the
future. We must provide these users a mechanism for navigating from a single commit
toward the root, allowing them to see a bigger picture.

\subsection{Core Model}

We root the tree with the merge made by Linus. Each merge may contain zero or more
children, which are other merges and commits. If a merge is cherry-picked or rebased
there will be no children. Commits have no children and are always leaf nodes. The
commits contains the metadata for the changes made. This metadata includes the files
changed, the lines added and removed from each file, the author, the date the commit
was merged into the merge that led to being merged into the kernel, and the date the
commit was authored, the diff, and the commit log. Merges contain less metadata,
only storing the author of the merge, the log, the commit date, and the author date.
The details of the model are outlined by German \cite{German2015}.

\section{Design}

In this section, we describe the progression of our tool design, showing the
benefits and drawbacks of each component.

\subsection{Command-line Shell}

Our initial intuition was to work with the commits and merges as directory trees
where the commits are files and the merges are directories. There are many
similarities between the tree structure of a file system and the tree structure of
the repository. Both trees are generally shallow, but very wide. File systems have
been studied for quite some time and there are many tools capable of visualizing and
working with the file structure.

The first tool we implemented was a small command-line shell (figure
\ref{fig:shell}) that allows users to navigate through the commits and merges rather
than files and directories. The shell design permitted a user to move through merges
like they would through directories with the \verb|cd| and \verb|ls| commands. Users
can use the \verb|cat| command to see the log preview, or the \verb|more| command to
see the preview and how many lines were added and removed from the files in the
commit.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/shell.png}
        \caption{Shell Design}
        \label{fig:shell}
\end{figure}

The shell does not meet our design goals. It only provides users with the ability to
navigate from the root. If there is a commit or merge that they would like to
investigate, they must find it with another tool and navigate through the tree to
find it. Furthermore, it does not aggregate information in the merges, so users must
manually aggregate the metadata. Another issue with this design is that it presents
the users with the commit hashes, which are quite long and unpleasant to remember
type into the shell, but also do not provide any quick explanation of what the
commit or merge may contain.

\subsection{Web-based Tool}

We replace the command-line interface with a web-based tool. Creating a web-based
tool enables users to use the system without having to install software or carry a
large database, making it more accessible, more easily maintainable, and platform
independent. A web-based tool can utilize more interactive means of navigation and
provide better explanations of commits and merges than the command-line interface.

Our tool uses the following mechanisms to reach our goals of better navigation and
better explanation of the changes.

\begin{itemize}
        \item Filter by search
        \item View files edited
        \item View modules
        \item Tree viewer
\end{itemize}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View}
        \label{fig:search}
\end{figure}

@@

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results.png}
        \caption{Search Results}
        \label{fig:results}
\end{figure}

In the search results, the user is presented with the one-line log message preview,
the author's name and email, the date the commit was authored, and the date the
commit was last committed. The commit date and author date will usually be the same;
however, if the commit has been rebased or cherry-picked into a merge that
eventually is merged into the kernel, the commit date and authored date will be
different. We can see this in the first and third entries in figure
\ref{fig:results}.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/message_view.png}
        \caption{Commit Message}
        \label{fig:message}
\end{figure}

Once a user selects a commit or merge to investigate, they are presented with a
tabbed pane allowing them to view the full commit log, the files edited, the modules
involved, and the tree views.

@@

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/file_view.png}
        \caption{Commit Files}
        \label{fig:files}
\end{figure}

The second tab is the files tab. This tab provides information on what files have
been edited, how many lines were added, and how many lines were removed in a given
commit. This functionality is similar to the other tools available. Our tree-based
design model allows us to extend this functionality to merges, which the other tools
are unable to show. When a file is edited multiple times from different commits
within a merge, we simply add the added and removed lines to the values already
stored.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules.png}
        \caption{Commit Modules}
        \label{fig:modules}
\end{figure}

@@

Finally, we have the tree view tab. The tree view is designed for providing easy
navigation within a tree and providing a clear topological view of the branch off
the kernel. We have experimented with various tree designs to find a design that
allows for easy navigation and visualization of both large and small trees. We
discuss the various tree designs in section \ref{treeview_section}.

\subsection{Tree Views} \label{treeview_section}

The tree view is what makes this tool unique to other tools, providing a full view
of the entire merge tree. We have experimented with various types of trees; list
trees are similar to the representation by many file system explorers,
Reingold-Tilford trees which clearly represent the tree structure, and bubble trees
which organize the data hierarchically by having the parent node contain the child
nodes.

\subsubsection{List Tree}

The list tree  viewer is in the form of nested lists, and is designed to more
closely model the tree hierarchy view found in file browsers. This tree only
contains the commits and merges that are within the subtrees of the current merge. A
commit will never have any items in this tree, as it is a leaf. To accompany the
tree, we include breadcrumbs at the top of the page to enable a user to navigate
both from the root to the leaves and from a leaf to the root. The last item in the
breadcrumb list is the current commit, the next item is the parent of the current
commit, and the first item is the root merge into the kernel.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
        \caption{List Tree}
        \label{fig:list_tree}
\end{figure}

\subsubsection{Reingold-Tilford Tree}

The Reingold-Tilford tree produces very nice results, creating an intuitive
representation of the tree. The representation has a clear notion of root and
leaves, and how to navigate in either direction.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
        \caption{Reingold-Tilford Tree}
        \label{fig:tree}
\end{figure}

Clicking on subtrees will cause them to minimize, only showing parts of the tree
that are of interest to the user. Furthermore, the user can zoom the tree by
scrolling the mouse wheel to see more of the topology of the tree, or to see the
details of a given part of the tree.

Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
thousands of commits and merges. While the tree is capable of producing a
visualization, it becomes much more difficult to understand.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
        \caption{Reingold-Tilford Tree zoomed out}
        \label{fig:zoomed_tree}
\end{figure}

\subsubsection{Bubble Tree}

The bubble tree provides a very clear picture of where a commit is located in the
hierarchy. The currently selected commit or merge is colored red, while non-selected
merges use a blue color and commits use white. The depth of the blue colouring is
dependent on how deep in the hierarchy that node is. The root is the lightest blue,
and the contained merges are progressively darker.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
        \caption{Bubble Tree}
        \label{fig:bubble_tree}
\end{figure}

The bubble tree doesn't have an implicit way of providing additional information,
placing any text near the nodes makes the tree impossible to read, so we include a
separate pane in the web page. When a user hovers over a node, the pane shows
additional information about the author and the commit message and a link to the
detailed page for that commit or merge. If the user clicks on a node, the tree will
zoom to that node and the information in the info pane will persist, enabling the
user to click the link. This tree provides an easy mechanism for users to navigate
from the root to the leaves and vice versa.

\section{Implementation Details}

\evan{Is this section necessary?}

We built a web-based tool to provide information about the kernel.  We chose to use
Nginx as the main web server, postgresql as the dbms, and Flask as the framework for
generating the dynamic content.

@@

Our tool provides mechanisms and visualizations that other tools are unable to
produce. Continuous mining of the repository\cite{German} provides more information
about how the repository is changing. It gives us the ability to see how commits are
being merged into the kernel, allowing us to more reliably build trees from the
snapshots of the repository.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
        \caption{Merge Dag View}
        \label{fig:dag_view}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.png}
        \caption{Merge Tree View}
        \label{fig:tree_view}
\end{figure}

@@

\subsection{Extendability}

The front-end uses asynchronous requests to gather the information for the trees and
tables. This enables third-parties to implement new front-ends, trying other
designs, though this interface may change in the future. The tree, file, and module
information is accessible through 
http://li.turingmachine.org/data/tree/JSON/\verb|<commit id>|,
http://li.turingmachine.org/data/files/JSON/\verb|<commit id>|,
http://li.turingmachine.org/data/modules/JSON/\verb|<commit id>|, respectively.

The response for the tree is a single object. This object is the root node of the
tree, and may contain an object of children objects in the \verb|children| field,
which are in the same format as the root object. The children object uses the commit
id as the key.

The tree node objects have the following fields:

\begin{tabular}{ccl}
        Field name & Data type & Description\\\hline
        cid & string & Commit id\\
        name & string & One-line log preview\\
        mlinus & string & Root merge commit id\\
        author & string & Author name and email\\
        mnext & string & Parent merge commit id\\
        children & object & Object of tree objects\\
\end{tabular}

The file responses contain only the files that the selected merge or commit works
with. The response is a single object in the form of a tree. This object is the root
of the tree, and represents the current merge or commit. If the current position in
the tree is an inner node, the response will contain the child nodes in the
``children'' field, otherwise ``children'' will be an empty list. The full response
format for the file response looks like:

\begin{itemize}
        \item ``cid'': string, commit id
        \item ``files'': list, list of tuples in the format
                \begin{enumerate}
                        \item Filename: string
                        \item Lines added: unsigned integer
                        \item Lines removed: unsigned integer
                \end{enumerate}
        \item ``mnext'': string, the commit id of the next merge toward the
                top-level merge
        \item ``children'': list, List of all child objects with the same
                fields. Null if no children.
\end{itemize}

\section{Future Work}

There are still many areas that can be improved before the full potential of our
model can be realized. Here, we outline various areas that still need more
attention.

\subsection{Files}

There is currently no functionality surrounding searching by filename. Use-case 2
may know what file they are editing and try to determine how it may work with the
other commits and merges in the module. It is possible that a third use-case may
arise, where the user wants to determine all commits that effect a given file. In
both cases, a bottom-to-top approach is used. With this functionality, implementing
the top-to-bottom search feature becomes trivial.

@@

\subsection{Authorship}

Our model can aggregate more information than what we have implemented. The
authorship information is important for licensing purposes, and we can show all
authors contributing to a merge and how many commits they contributed to the merge.
We could go further, providing information about what files they edited, how many
lines they added and removed, and more.

\subsection{Diff generation}

Due to the construction of the tree, it is possible to generate a diff of a given
file in a merge. This would remove areas that overlap, and provide users with a
single patch at a merge rather than attempting to combine all of the diff files from
the leaves.

\subsection{Evaluation}

At this time, we have no evidence that our tool is able to improve the work-flow of
maintainers. We believe that the tool is able to improve the work-flow and
performance of maintainers because it provides cleaner mechanisms of visualizing and
presenting the commit information. It is able to provide more relevant information,
while removing information that is irrelevant to a given module or set of merges.
